### 2018-03-11 paper/3-compute.tex

...
% \begin{figure}
% \begin{minted}{haskell}
% type Compute c k v = @\std{forall}@ f. c f => (k -> f v) -> k -> Maybe (f v)
% \end{minted}
% \caption{Compute abstractions}\label{fig-compute}
% \end{figure}
% Consider abstractions in Fig.~\ref{fig-compute}.

>>>>

We describe build tasks and their dependencies using functions of type
\hs{Compute}:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
type Compute c k v = @\std{forall}@ f. c f => (k -> f v) -> k -> Maybe (f v)
\end{minted}
\vspace{1mm}

\noindent
A compute takes two arguments: \hs{fetch :: k -> f v} and \hs{key :: k}, and
returns either \hs{Nothing} to indicate that the \hs{key} is an input and hence
cannot be computed, or \hs{Just value} obtained by \hs{fetch}ing the
dependencies of the \hs{key} and performing the required computation. A compute
must be polymorphic in \hs{f} but may request certain computational
power\footnote{Readers familiar with \emph{lenses} might recognise a familiar
pattern. We discuss this in~\S\ref{sec-related}.} by specifying the
constraint~\hs{c}. Let us look at an example by recalling the following
spreadsheet from~\S\ref{sec-background-excel}:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{bash}
A1 = 10
A2 = 20
B1 = A1 + A2
\end{minted}
\vspace{1mm}

\noindent
We can express it as an \emph{applicative compute} as follows:
\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
data Cell = A1 | A2 | B1 deriving Eq

add :: Compute Applicative Cell Integer
add fetch key | key /= B1 = Nothing
              | otherwise = Just $ (+) <$> fetch A1 <*> fetch A2
\end{minted}
\vspace{1mm}

\noindent
Cells \cmd{A1} and \cmd{A2} are inputs and \hs{add} returns \hs{Nothing} to
indicate this. The only task is associated with the cell \cmd{B1}: fetch the
values of \cmd{A1} and \cmd{A2} and apply the pure function~\hs{(+)} to them
using the methods of \hs{Applicative} type class.

%\hs{($) :: (a -> b) -> f a -> f b} and \hs{(<*>) :: f (a -> b) -> f a -> f b}

% \begin{figure}
% \begin{subfigure}[b]{\linewidth}
% \begin{minted}{haskell}
% type Compute c k v = @\std{forall}@ f. c f => (k -> f v) -> k -> Maybe (f v)
% \end{minted}
% \caption{Compute}
% \end{subfigure}
% ~\\
% \begin{subfigure}[b]{\linewidth}
% \begin{minted}{haskell}
% dependency :: Compute Functor k v -> k -> Maybe k
% dependency compute = fmap getConst . compute Const
% \end{minted}
% \caption{Functorial compute}
% \end{subfigure}
% \caption{Compute abstractions}\label{fig-compute}
% \end{figure}

% Consider abstractions in Fig.~\ref{fig-compute}.
